\textsc{\wobble: a data-driven method for precision radial velocities}

### 2019-07-26 paper/paper.tex
edits and new script

The stellar spectrum contribution is: 
\begin{equation}
f_{\star,n} = P(\xi_n, \xi_{\star}, v_{obs,n}) \cdot \mu_{\star},
\label{eqn:star}
\end{equation}
where $\mu_{\star}$ is a spectral template of log-fluxes and 
$\xi_{\star}$ is the corresponding vector of template log-wavelengths. 
The exact values of $\xi_{\star}$ can be chosen somewhat arbitrarily under the conditions that the grid has uniform spacing $\Delta\xi_{\star}$, covers the entire wavelength range of the data, and is over-sampled with regards to the observed spectrum's wavelength grid.
$P$ is a linear operator whose function is to apply a Doppler shift by observed velocity $v_{obs,n}$ and
interpolate $\mu_{\star}$ from the $\xi_{\star}$ template grid to the $\xi_n$ data grid.
Each entry of the $P$ matrix can be defined by a sum of weighted indicator functions as follows:
\begin{equation}
\begin{split}
 P_{i,j} =  &\ \Big(\frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}}\Big) \mathbb{1}\big(0 < \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}} < 1\big) \\
  &+ \Big(1 - \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}}\Big) \mathbb{1}\big({-1} < \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}} < 0\big), \\
 \end{split}
\end{equation}
where $\xi'_{\star}$ is the Doppler-shifted template grid:
\begin{equation}
 \xi'_{\star}(v) = \xi_{\star} + D(v) \equiv \xi_{\star} + \frac{1}{2} \ln \left(\frac{1 - v/c}{1 + v/c}\right).
\end{equation}
Because $P$ is quite sparse, in practice we do not instantiate the full matrix when performing calculations. 
At this point, it is also useful to note that $P$ could, in principle, encode an instrumental line spread function (\LSF). 
We return to this point in Section \label{s:future}.

@@

Another way a less well-calibrated instrument may violate \wobble's assumptions is in the non-negligible variability of all spectral lines due to changes in the instrumental line spread function (\LSF). 
In principle, if the changes to the \LSF were well-understood they could be hard-coded into linear operators like the $P$ operator used in Equation \ref{eqn:star}, leaving the essential equations of \wobble largely unchanged. 
\added{In this case the \LSF operator should be applied to both the stellar and telluric component models after combining them.} 
Realistically speaking, if \LSF variations are present they are probably not sufficiently well-understood to take this approach; instead, we might want to fit these variations as a data-driven model component. 
This leads us into the territory of making more fundamental changes to the \wobble model.

@@

\acknowledgements{\added{We gratefully acknowledge the referee, Ian Czekala, for a thoughtful and constructive review.}
It is a pleasure to acknowledge Jacob L. Bean, John M. Brewer, Heather Knutson, Timothy Morton, Melissa Ness, Andreas Seifahrt, and Julian St\"{u}rmer for helpful discussions. 
We are grateful to Hans-Walter Rix and the Max-Planck-Institut f\"ur Astronomie for their hospitality while developing key parts of this project. Special thanks go to Geert Barentsen, Michael Gully-Santiago, Christina Hedges, Susan Mullaly, and Z\'{e} Vin\'{i}cius for their vital contribution to the most important part of \wobble's documentation (the pronunciation guide).

>>>>

The stellar spectrum contribution is: 
\begin{equation}
f_{\star,n} = P(\xi_n, \xi_{\star}, v_{obs,n}) \cdot \mu_{\star},
\label{eqn:star}
\end{equation}
where $\mu_{\star}$ is a spectral template of log-fluxes and 
$\xi_{\star}$ is the corresponding vector of template log-wavelengths. 
The exact values of $\xi_{\star}$ can be chosen somewhat arbitrarily under the conditions that the grid has uniform spacing $\Delta\xi_{\star}$, covers the entire wavelength range of the data, and is over-sampled with regards to the observed spectrum's wavelength grid.
$P$ is a linear operator whose function is to apply a Doppler shift by observed velocity $v_{obs,n}$ and
interpolate $\mu_{\star}$ from the $\xi_{\star}$ template grid to the $\xi_n$ data grid.
Each entry of the $P$ matrix can be defined by a sum of weighted indicator functions as follows:
\begin{equation}
\begin{split}
 P_{i,j} =  &\ \Big(\frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}}\Big) \mathbb{1}\big(0 < \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}} < 1\big) \\
  &+ \Big(1 - \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}}\Big) \mathbb{1}\big({-1} < \frac{\xi_{n,i} - \xi'_{\star,j}}{\Delta\xi_{\star}} < 0\big), \\
 \end{split}
\end{equation}
where $\xi'_{\star}$ is the Doppler-shifted template grid:
\begin{equation}
 \xi'_{\star}(v) = \xi_{\star} + D(v) \equiv \xi_{\star} + \frac{1}{2} \ln \left(\frac{1 - v/c}{1 + v/c}\right).
\end{equation}
Because $P$ is quite sparse, in practice we do not instantiate the full matrix when performing calculations. 
At this point, it is also useful to note that $P$ could, in principle, encode an instrumental line spread function (\LSF). 
We return to this point in Section \ref{s:future}.

@@

Another way a less well-calibrated instrument may violate \wobble's assumptions is in the non-negligible variability of all spectral lines due to changes in the instrumental line spread function (\LSF). 
In principle, if the changes to the \LSF were well-understood they could be hard-coded into linear operators like the $P$ operator used in Equation \ref{eqn:star}, leaving the essential equations of \wobble largely unchanged. 
\added{In this case the \LSF operator should be applied to both the stellar and telluric component models after combining them. Indeed, technically the current \wobble model does not deal correctly with line broadening: the templates we infer for star and tellurics are post-broadening spectra, but this goes against the correct order of operations if line broadening is dominated by instrumental effects that are applied to the \textit{combined} spectra. We have assumed that this effect is negligible for the purposes of this work, but it should certainly be considered more carefully in the future.}

Realistically speaking, if \LSF variations are present they are probably not sufficiently well-understood to \replaced{take this approach}{take a prescriptive approach to their behavior and hard-code them into the form of the $P$ operator}; instead, we might want to fit these variations \replaced{as a data-driven model component}{using a data-driven approach, which would make the \wobble model significantly non-linear}. 
This leads us into the territory of making more fundamental changes to the \wobble model.

@@

\acknowledgements{\added{We gratefully acknowledge the referee, Ian Czekala, for a thoughtful and constructive review.}
It is a pleasure to acknowledge Jacob L. Bean, John M. Brewer, Heather Knutson, Timothy Morton, Melissa Ness, Andreas Seifahrt, Julian St\"{u}rmer, \added{and Sharon Xuesong Wang} for helpful discussions. 
We are grateful to Hans-Walter Rix and the Max-Planck-Institut f\"ur Astronomie for their hospitality while developing key parts of this project. Special thanks go to Geert Barentsen, Michael Gully-Santiago, Christina Hedges, Susan Mullaly, and Z\'{e} Vin\'{i}cius for their vital contribution to the most important part of \wobble's documentation (the pronunciation guide).
